A Detection of the Environmental Dependence of the Sizes and Stellar Haloes of Massive Central Galaxies

### 2017-07-31 draft/paper2_v4.tex
Edited discussion

    We use ${\sim}100$ deg$^2$ of deep ($>28.5$\sb{} in $i$-band), high-quality 
    (median 0.6\asec seeing) imaging data from the Hyper Suprime-Cam (HSC) survey 
    to investigate the halo mass dependence of the surface mass density profiles 
    and outer stellar envelopes of massive galaxies. 
    The depth of the HSC survey reaches ${\sim}4$ magnitudes fainter than SDSS and 
    enables us to directly trace stellar mass distributions to 100 kpc without 
    requiring stacking.  
    We conclusively show that at fixed stellar mass, the stellar profiles of massive 
    galaxies depend on the masses of the their dark matter halos. 
    On average, massive central galaxies in more massive halos have shallower inner 
    stellar mass density profiles (within ${\sim}10$-$20$ kpc) and more prominent 
    outer envelopes. 
    These differences translate into a clear environmental dependence of the 
    mass-size relation: central galaxies in more massive halos have larger 
    $R_{\mathrm{50}}$ at fixed \mtot{}. 
    We further find that the ellipticity of the outer stellar halo also increases 
    with halo mass. 
    Our results are broadly consistent with the two phase formation scenario of 
    massive galaxies in which outer envelopes are built up at late times from the 
    accretion of small systems.

@@

    \begin{itemize}
        \item \alexie{discuss here}. Massive central galaxies living in more and 
            less massive haloes display similar ellipticity profiles at fixed \mtot{}. 
            There is hint that suggests the central galaxies from more massive 
            haloes become slightly more elongated at $R>50$ kpc than the ones 
            in less massive haloes with similar \mtot{}. 
            However, the uncertainties in the current ellipticity profiles make it 
            not conclusive.

@@

    %Using the deep images from the HSC survey, we show that the stellar mass 
    %distributions of massive central galaxies at $0.3 < z < 0.5$ depend on their 
    %environment or \mhalo{}. 
    %Here we briefly discuss the physical implications and potential caveats of this 
    %result.

%% ------------------------------------------------------------------------------------ %% 
\subsection{The Role of Environment in the ``Two--Phase'' Formation Scenario}

\alexie{This section does not have a good logical flow and needs more work and clarity. Ideas are ``choppy''}.

    Our results are broadly consistent with the ``two-phase'' formation theory for massive central galaxies. According to this scenario, intense dissipative processes at $z > 2$ are responsible for 
    the formation of \textit{in situ} stars in massive central galaxies. In the second phase, star-formation is quenched which limits in-situ star formation and growth becomes dominated by non-dissipative mergers of satellite galaxies. Dry minor mergers are efficient at depositing \textit{ex situ} stars in the outskirts of massive galaxies (e.g. \citealt{Oogi2013, Bedorf2013}). Because central galaxies in massive haloes should experience more mergers compared to those in less massive halos, we expect the stellar mass profiles (and sizes) of massive central galaxies to exhibit dependencies on host halo mass.

    \citet{Shankar2014} studied the environment dependence of galaxy size using 
    a series of hierarchical semi-analytic models. They predict that at fixed stellar mass, the median size of central galaxies should increase strongly with halo mass. The major-merger rate increases slightly with halo mass (mass ratio $>1$:3; e.g. \citealt{Hirschmann2013}). The minor-merger rate, however, depends strongly on the dynamical friction time. If the dynamical friction time is short (e.g. \citealt{Newman2012}), then centrals of more massive haloes can have up to four times more minor mergers (1:100--1:3) \alexie{Compare to what? Need to give the halo mass numbers here}. This increase in the minor-merger rate as a function of halo mass leads to a halo-mass dependence in the mass-size relation. The prediction from \citet{Shankar2014} have been confirmed by \citet{Yoon2017} using the semi-analytic model from \citet{Guo2011}.  Our results are broadly consistent with these predictions. Indeed, we find that the outer profiles of massive galaxies ($r>50$ kpc) are the regions that display the strongest variations as a function of halo mass.

    %Whether or not level of difference that we observe is fully consistent with predictions from semi-analytic or hydrodynamic models of galaxy formation is work in progress that will be presented in a follow-up paper.

    More importantly, our results show that the inner region of massive central 
    galaxies also play a role here. 
    Both the \mden{} profiles of \mtot{}--matched samples and the \mtot{}--\minn{}
    plane suggest that, at fixed \mtot{}, centrals of more massive haloes have 
    slightly shallower inner \mden{} slope and lower fraction of \mtot{} within the 
    inner 10 kpc.
    Minor mergers can not explain such difference as they should not alter the inner 
    \mden{} profiles. 
    This is also not caused by the choice of outer aperture to estimate ``total''
    \mstar{}: assuming that centrals in more massive haloes have more extended 
    \mden{} profiles, integration out to larger aperture will make the differences 
    between \rbcg{} and \nbcg{} galaxies on the \mtot{}--\minn{} plane more 
    significant.

    Observations of compact quiescent galaxies at high-$z$ seem to indicate higher 
    \mden{} in their central regions than their descendants. 
    Therefore it is necessary to rely on certain processes to reduce the central 
    \mden{} (e.g. strong adiabatic expansion induced by powerful AGN feedback; 
    \citealt{Fan2008}) to match the inner \mden{} profiles of massive central galaxies 
    at lower redshift.
    Coalesce of super-massive black hoes (SMBHs) at the end of major merger, which 
    is common in the assembly history of massive galaxies, can also flatten the 
    central \mden{} profile via efficient scattering effect 
    (e.g. \citealt{Milosavljevi2002}).
    On the right side of Fig \ref{fig:scaling_relation}, we find a few \rbcg{} 
    galaxies with very low \minn{} at high \mtot{} end that are apparently not bothered 
    by problematic photometry or exceptionally bad seeing.  
    They may be similar to the recently discovered massive BCGs with very large 
    depleted cores (a few kpc; e.g. \citealt{Postman2012, LopezCruz2014, Thomas2016,
    Bonfini2016}) resulting from SMBH mergers.
    The impacts of these processes, and their dependence on \mhalo{} is an interesting 
    topic to investigate in the future.

@@

    \citet{Cerbrian2014} find that ETGs with \logms{}$>11.5$ are slightly larger in 
    more massive haloes, but they also shows that this trend is reversed 
    at lower \mstar{}. \citet{Yoon2017} present a careful study of a large sample of $z\sim0.1$ SDSS ETGs using  
    a non-parametric method and recover a clear environment dependence of mass-size 
    relation at \logms{}$>11.2$. They conclude that, at the same \mstar{}, massive ETGs from dense environment 
    are 20--40\% larger compared to those in underdense environment -- these results are in broad agreement with those presented here.
    \alexie{Need a sentence abbout the new Mike Hudson paper here}

    While there now appears to be an emerging consensus that the high mass end of the mass-size relation does vary with halo mass, the situation at higher redshift is more complicated. Certain studies find an environment dependence of the mass--size relation for massive 
    quiescent or early-type galaxies that is qualitatively consistent with low redshift studies (e.g. \citealt{Papovich2012, Lani2013, Strazzullo2013, 
    Delaye2014}). However, there are also observations that suggest otherwise (e.g. 
    \citealt{Rettura2010, Raichoor2012, Kelkar2015, Allen2015}). These differences might be explained by varying sample sizes,  special environments
    (high-$z$ massive cluster), or the impact of massive star-forming galaxies. However, these yet un-reconciled differences, also underscore the  intrinsic difficulty of measuring stellar masses and sizes 
    for high-$z$ galaxies (e.g. \cite{Price2017}). Instead of using the mass--size relation to compare the high and the low redshift universe, a more robust approach would be to directly compare rest-frame surface brightness 
    or stellar mass density profiles (e.g. \citealt{Szomoru2012, Patel2013, Buitrago2017, Hill2017}).

    \subsection{Towards Consistent Size Definitions}

          Semi-analytic models and hydrodynamic simulations often have difficulties reproducing the  mass--size relation of massive galaxies. Galaxy sizes are sensitive to many different physical processes (star-formation, feedback, mergers) and matching the galaxy stellar-mass function does not automatically guarantee a match to the mass-size relation. Furthermore, while some effort has been made to use consistent size definitions (cite Bahamas paper), more often than not, comparisons of the mass-size relation often do not use consistent size definitions, or only perform crude size conversions (give an example here? 3d radius and 2 radius). For examples, observers often quote ``sizes" which are the effective radius of an elliptical sersic fit. This is a 2d size  corresponding to the major axis of the ellipse that contains 50\% of the light. Simulations on the other hand, often employ sizes that correspond to xxx. In this paper, we quote masses measured within elliptical apertures of fixed physical size. We argue that this approach will allow for a more straightforward comparison between observations and theoretical predictions. In follow-up work, we will present comparisons between our observations and predictions from hydrodynamic simulations using the  using the \mtot{}--\minn{} plane instead of the mass-size relation.

>>>>

    We use ${\sim}100$ deg$^2$ of deep ($>28.5$\sb{} in $i$-band), high-quality 
    (median 0.6\asec seeing) imaging data from the Hyper Suprime-Cam (HSC) survey 
    to investigate the halo mass dependence of the surface mass density profiles 
    and outer stellar envelopes of massive galaxies. 
    The depth of the HSC survey reaches ${\sim}4$ magnitudes fainter than SDSS and 
    enables us to directly trace stellar mass distributions to 100 kpc without 
    requiring stacking.  
    We conclusively show that at fixed stellar mass, the stellar profiles of massive 
    galaxies depend on the masses of the their dark matter halos. 
    On average, massive central galaxies in more massive halos have shallower inner 
    stellar mass density profiles (within ${\sim}10$-$20$ kpc) and more prominent 
    outer envelopes. 
    These differences translate into a clear environmental dependence of the 
    mass-size relation: central galaxies in more massive halos have larger 
    $R_{\mathrm{50}}$ at fixed \mtot{}.  
    Our results are broadly consistent with the two phase formation scenario of 
    massive galaxies in which outer envelopes are built up at late times from the 
    accretion of small systems.

@@

    \begin{itemize}
        \item Within $R>50$ kpc, we find that the ellipticity profiles of massive 
            central galaxies do not depend on halo mass at fixed \mtot{}.

@@

%% ------------------------------------------------------------------------------------ %% 
\subsection{The Role of Environment in the Two-Phase Formation Scenario}

    Using the deep images from the HSC survey, we show that the stellar mass 
    distributions of massive central galaxies at $0.3 < z < 0.5$ depend on their 
    environment or \mhalo{}. 
    At fixed \mtot{} or \minn{}, we find that the central galaxies from more massive 
    haloes on average have larger half--light radius and host more prominent outer 
    stellar haloes.

    These results are broadly consistent with the ``two-phase'' formation theory for 
    massive central galaxies. 
    According to this scenario, intense dissipative processes at $z > 2$ are responsible 
    for the formation of \textit{in situ} stars in massive central galaxies. 
    In the second phase, star-formation is quenched which limits in-situ star formation 
    and growth becomes dominated by non-dissipative mergers of satellite galaxies. 
    Dry minor mergers are efficient at depositing \textit{ex situ} stars in the 
    outskirts of massive galaxies (e.g. \citealt{Oogi2013, Bedorf2013}). 
    Because central galaxies in massive haloes should experience more mergers compared 
    to those in less massive halos, we expect the stellar mass profiles (and sizes) of 
    massive central galaxies to exhibit dependencies on host halo mass.

    \citet{Shankar2014} studied the environment dependence of galaxy size using 
    a series of hierarchical semi-analytic models. 
    They predict that at fixed stellar mass, the median size of central galaxies 
    should increase strongly with halo mass. 
    The major-merger rate only increases slightly with halo mass (mass ratio $>1$:3; 
    e.g. \citealt{Hirschmann2013}). 
    The minor-merger rate, however, depends strongly on the dynamical friction time. 
    If the dynamical friction time is short (e.g. \citealt{Newman2012}), the
    massive central galaxies with $11.5 <$\logms{}$<12.0$ from more massive haloes 
    (\logmh{}$>14.0$) can have up to four times more minor mergers (1:100--1:3)
    comparing to the ones in less massive halos at similar stellar mass. 
    This increase in the minor-merger rate as a function of halo mass leads to a 
    halo-mass dependence in the mass-size relation. 
    The prediction from \citet{Shankar2014} have been confirmed by \citet{Yoon2017} 
    using the semi-analytic model from \citet{Guo2011}.  
    Our results are broadly consistent with these predictions. 
    Indeed, we find that the outer profiles of massive galaxies ($r>50$ kpc) are the 
    regions that display the strongest variations as a function of halo mass.

%% ------------------------------------------------------------------------------------ %% 
\subsection{Environmental Dependence of the Inner Regions of Massive Galaxies}

    At the same time, from the comparison of median \mden{} profiles of the 
    \mtot{}--matched samples and from the \mtot{}--\minn{} plane, our results also 
    show that the centrals of more massive haloes have slightly shallower inner 
    \mden{} slope and lower fraction of \mtot{} within the inner 10 kpc.
    It seems that the environmental dependence of stellar mass distributions in 
    massive central galaxies also affects their central region.
    Unlike the differences in the outer stellar halo, it is difficult to use minor 
    mergers to explain such difference as they should not alter the inner \mden{} 
    profiles. 
    It is also not caused by the choice of outer aperture to estimate ``total''
    \mstar{}: assuming that centrals in more massive haloes have more extended 
    \mden{} profiles, integration out to larger aperture will make the differences 
    between \rbcg{} and \nbcg{} galaxies on the \mtot{}--\minn{} plane more 
    significant.

    Several physicals processes may help shape the stellar mass distributions in the 
    inner region. 
    For instance, strong adiabatic expansion induced by powerful AGN feedback can 
    lead to more flattened inner \mden{} profile \citealt{Fan2008}. 
    The violent relaxation process involved in the major dry merger can reshape 
    stellar mass distributions in the inner region. 
    In addition, the coalesce of super-massive black hoes (SMBHs) at the end of 
    major merger, which is common in the assembly history of massive galaxies, can 
    also flatten the central \mden{} profile via efficient scattering effect 
    (e.g. \citealt{Milosavljevi2002}).
    On the right side of Fig \ref{fig:scaling_relation}, we find a few \rbcg{} 
    galaxies with very low \minn{} at high \mtot{} end that are apparently not bothered 
    by problematic photometry or exceptionally bad seeing.  
    They may be similar to the recently discovered massive BCGs with very large 
    depleted cores (a few kpc; e.g. \citealt{Postman2012, LopezCruz2014, Thomas2016,
    Bonfini2016}) resulting from SMBH mergers.
    The impacts of these processes, and their dependence on \mhalo{} is an interesting 
    topic to investigate in the future through comparisons with both observations 
    at high redshift and numerical simulations.

@@

    \citet{Cerbrian2014} find that ETGs with \logms{}$>11.5$ are slightly larger in 
    more massive haloes, but they also shows that this trend is reversed 
    at lower \mstar{}. 
    \citet{Yoon2017} present a careful study of a large sample of $z\sim0.1$ SDSS ETGs 
    using a non-parametric method and recover a clear environment dependence of 
    mass-size relation at \logms{}$>11.2$. 
    They conclude that, at the same \mstar{}, massive ETGs from dense environment 
    are 20--40\% larger compared to those in underdense environment -- these results are 
    in broad agreement with those presented here.
    Recently, \citet{Charlton2017} uses single-\ser{} model for galaxies at 
    $0.2 < z < 0.8$ in the CFHTLenS (\citealt{Heymans2012}) and their galaxy--galaxy 
    lensing measurements to show that larger galaxies tend to live in more massive 
    dark matter haloes.

    While there now appears to be an emerging consensus at low redshift that the 
    high mass end of the mass-size relation does vary with halo mass, the situation at 
    higher redshift is more complicated. 
    Certain studies find an environment dependence of the mass--size relation for 
    massive quiescent or early-type galaxies that is qualitatively consistent with low 
    redshift studies (e.g. \citealt{Papovich2012, Lani2013, Strazzullo2013, 
    Delaye2014}). 
    However, there are also observations that suggest otherwise (e.g. 
    \citealt{Rettura2010, Raichoor2012, Kelkar2015, Allen2015}). 
    These differences might be explained by varying sample sizes, special environments
    (high-$z$ massive cluster), or the impact of massive star-forming galaxies. 
    However, these yet un-reconciled differences, also underscore the intrinsic 
    difficulty of measuring stellar masses and sizes for high-$z$ galaxies 
    (e.g. \cite{Price2017}). 
    Instead of using the mass--size relation to compare the high and the low redshift 
    universe, a more robust approach would be to directly compare rest-frame surface 
    brightness or stellar mass density profiles (e.g. \citealt{Szomoru2012, Patel2013,
    Buitrago2017, Hill2017}).

%% ------------------------------------------------------------------------------------ %% 
\subsection{Towards Consistent Size Definitions}

    Until recently, semi-analytic models and hydrodynamic simulations still have 
    difficulties reproducing the mass--size relation of massive galaxies. 
    Galaxy sizes are sensitive to many different physical processes (star-formation, 
    feedback, mergers) and matching the galaxy stellar-mass function does not 
    automatically guarantee a match to the mass-size relation. 
    Furthermore, while some efforts have been made to use consistent size definitions 
    (\citealt{McCarthy2017}), more often than not, comparisons of the mass-size relation 
    often do not use consistent size definitions, or only perform crude size conversions 
    (e.g. 3--D radius in simulation versus 2--D projected radius in observation; 
    \citealt{Genel2017}). 
    Observers often quote ``sizes'' corresponding to the half--light radius along the 
    major axis using 2--D projected images. 
    Simulations on the other hand, often employ sizes that correspond to the 3--D 
    aperture half--mass radius (e.g. \citealt{Price2017}).

    As emphasized earlier, definitions of galaxy ``size'' in observations are also not 
    always consistent. 
    Measurement of ``size'' often show dependence on image quality (e.g. seeing, 
    imaging depth), filter, and the adopted method. 
    Although elliptical single--\ser{} model is widely adopted in measuring the 
    size of galaxies of different types and at different redshift, it sometimes leads
    to biased results as it is no longer appropriate model for certain type of galaxy. 
    Size derived from 1-D curve--of--growth is more model independent and is shown to 
    be useful in revealing environmental dependence of mass--size relation 
    (e.g. \citealt{Yoon2017} and this work), but it does not take the PSF convolution 
    effect into account and depends on imaging depth and background subtraction.

    In this paper, we quote stellar masses measured within elliptical apertures of 
    fixed physical size. 
    We argue that this approach will allow for a more straightforward comparison 
    between observations and theoretical predictions. 
    In follow-up work, we will present comparisons between our observations and 
    predictions from hydrodynamic simulations using the using the \mtot{}--\minn{} 
    plane instead of the mass--size relation.
